---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First of all we clean the workspace as you can read on https://support.rstudio.com/hc/en-us/articles/200711843-Working-Directories-and-Workspaces and unzip the data file, which create a ```activity.csv``` file. If ```activity.csv``` exists it will be overwritten.

The scripts stops if no file ```activity.zip``` was found in working directory or it couldn't be unzipped.

```{r echo=FALSE, "Seeting working directory"}
setwd("T:/Formacion/Data Scientist/05 Reproducible research/Assignment 1/RepData_PeerAssessment1/")
```

```{r "Clean workspace and unzip data file"}
rm(list = ls())
zipfile <- "activity.zip"
if (!file.exists(zipfile)) {
    stop ("No activity.zip zipped data file found on working directory")
}
tryCatch({
    unzip("activity.zip", overwrite = TRUE)
}, warning = function(war) {
    stop("Some warnings occur extracting file. Process has stopped.")
}, error = function(err) {
    stop("Error unzipping data file. Process has stopped.")
})
```

Once we have tha CSV file, we import its data into activity data.frame.

```{r "Load CSV data file into activity data.frame"}
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

First we calculate total steps by day using ```aggregate``` function and plot an histogram with resulting data.frame.

```{r "Steps by day"}
steps_by_day <- aggregate(steps ~ date, activity, sum)
hist(steps_by_day$steps, main = "Steps taken by day", xlab = "Num. of steps")
```

And now calculate mean and median

```{r "Mean and media steps by day"}
mean(steps_by_day$steps)
median(steps_by_day$steps)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
